---
title: "R markdown example"
author: "A. Paxton"
output: html_document
---

This is an example R markdown file, created as part of the living documents and
reproducible reports workshop.The GitHub repository for this workshop is
available here: https://github.com/a-paxton/living-documents

In this R markdown file, we'll analyze the toy data provided in the repository,
which is the same data that we generated with our example Jupyter notebook
(`/jupyter-notebooks/jupyter_notebook-fake_experiment_data_generator.ipynb`).

**Created by:** A. Paxton (University of Connecticut)

**Date last modified**: 10 July 2020

***

# Preliminaries

First, we'll clear our workspace and make sure that we have all of the packages
we need.

```{r preliminaries}

# clear the workspace
rm(list=ls())

# install and/or load packages specified in another file
source('r_library_installation.r')

```

The `source()` function allows us to execute code included in another file of R
code. It's a handy function; if you use it, remember to use the relative or
absolute file path to it. (In other words, it will only look in the specific
directory for it, so if you don't include a different directory, it will only
look or a file in the working directory.)

Once that's done, we'll load in the toy dataset from this repository  (generated
in the `jupyter-notebooks/jupyter_notebook-fake_experiment_data_generator.ipynb`
Jupyter notebook).

```{r load-data}

# load in the dataset
experiment_df = read.table('../data/simulated_experiment_data.csv',
                           sep = ',',
                           header = TRUE)

# convert factor variables to factor
experiment_df = experiment_df %>% ungroup() %>%
  mutate(condition = as.factor(condition))

```

***

# Data checks

Before we get moving, let's make sure that our data look like we expect them to
look.

```{r print-data-header}

# check out the first few rows of data
head(experiment_df)

```

We can also get some summary stats on our dataset.

```{r print-summary-stats}

# get some basic information about each variable
summary(experiment_df)

```

## Plotting

We're always encouraged to take a look at our data as part of our data checks. R
markdown allows you to add those useful plots right into your living documents
and reproducible reports so that everyone gets to see how lovely your data are.

You can make your plots pretty basic:

```{r plot-negative-similarity-distribution}

# how do the negative similarity scores look?
qplot(negative_word_similarity,
      data = experiment_df,
      geom = "histogram",
      bins = 20)

```

Or you can make them more informative:

```{r plot-positive-similarity-distribution}

# how about positive similarity scores by condition?
ggplot(data = experiment_df,
       aes(x = positive_word_similarity,
           color = condition)) +
  geom_density() +
  labs(x = 'Positive word similarity',
       y = 'Density',
       color = 'Condition') +
  scale_color_manual(values=c('red','blue'),
                     breaks = c(0, 1),
                    labels=c('Commiseration', 'Celebration'))
```

Making your plots polished at first might take a bit longer, but providing as
much information as possible to yourself while creating a living document allows
you to more easily refresh your memory after a break *and* have to do less
cleanup when you want to share it as a reproducible report.

***

# Participant information

We can use living documents to get to know our data and to easily provide the
kind of information we'd need to share in an eventual manuscript. For instance,
you can easily re-calculate the total number of participants (or, in our case,
dyads) and their distribution across conditions.

## How many total dyads did we have?

```{r unique-dyad-IDs}

# figure out what unique dyad IDs we have
unique_dyads = experiment_df %>% ungroup() %>%
  select(dyad_ID) %>%
  distinct()

# display the unique dyad IDs
unique_dyads$dyad_ID

```

Instead of just displaying numbers, we can also get R to print it out in text
for us.

```{r unique-dyad-count}

# how many dyads did we have?
paste0("Total dyads: ",length(unique_dyads$dyad_ID))

```

We can also call variable printouts in the text of our reproducible reports.
For example, we have `r length(unique_dyads$dyad_ID)` total dyads in our sample.

(Note that the code will continue to show up in the RStudio window, but it will
be replaced by the correct value in the knitted file.)

## How many dyads were in each condition?

```{r condition-information}

# break down the data by condition
condition_dyads = experiment_df %>% ungroup() %>%
  select(condition, dyad_ID) %>%
  distinct()

# display condition information
condition_dyads

```

```{r condition-dyad-counts}

# get total condition counts
condition_counts = condition_dyads %>% ungroup() %>%
  group_by(condition) %>%
  summarise(dyad_counts = n())

# display it
condition_counts

```

Or, again, we can display it a bit more nicely:

```{r print-conditions-nicely}

# get dyads in each conditions
commiseration_dyads = condition_counts %>% ungroup() %>%
  filter(condition==0) %>%
  .$dyad_counts
celebration_dyads = condition_counts %>% ungroup() %>%
  filter(condition==1) %>%
  .$dyad_counts

# print each one
paste0('Total dyads in commiseration condition: ',
       commiseration_dyads)
paste0('Total dyads in celebration condition: ',
       celebration_dyads)


```


***

# Data analysis

We can also get help information from each cell. If we want to perform an ANOVA,
for example, we can get some help text on how to use the function:

```{r check-anova-call}

?anova

```

It's a good thing we checked -- because that's not the right call! It looks like
we need this one:

```{r check-aov-call}

?aov

```

Now that we've got the call down, let's run our model!

```{r data-analysis}

summary(aov(rapport ~ negative_word_similarity * positive_word_similarity * condition,
            data = experiment_df))

```

It looks like we've got two significant main effects: Using more similar amounts
of negative words and being involved in the commiseration condition are both
associated with higher rapport.

